---
title: "Financial Analytics Assignment 2 - Jasmeet Singh Sandhu"
output:
  html_document:
    df_print: paged
date: "2023-01-23"
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(quantmod)
library(PerformanceAnalytics)
```

The names of the selected companies are in the file Industrial_Health_Names.csv The period of observation is from="2014-7-1", to="2015-7-1"

For the sector indices use SPDR XLV (health care sector) and XLI (industrial sector) For the broad market index use SPY For the risk free rate use Fed Funds effective rate

Note that it may not be possible to find interpretation of PCA factors in terms of real assets or indices. In such cases it is possible to use PCA factors without interpretation

```{r, warning=FALSE, include = FALSE}

# Read list of stock names
datapath <- 'C:\\Users\\jasme\\Downloads\\UChicago\\Quarter2\\Financial Analytics\\Week2\\'
SP500.Industrials.Health <- read.table(paste(datapath,'assignments_Fonancial Analytics (32001)_Lecture 2_Industrials_Health_Names.csv',sep = '/'), sep = ',', header = FALSE)
SP500.Industrials.Health.names <- as.character(SP500.Industrials.Health[,1])
SP500.Industrials.Health.names

getSymbols(SP500.Industrials.Health.names,from="2014-7-1",to="2015-7-1")
getSymbols("XLV", from="2014-7-1", to = "2015-7-1")
getSymbols("XLI", from="2014-7-1" ,to = "2015-7-1")
getSymbols("SPY", from="2014-7-1", to = "2015-7-1")
getSymbols("DFF", src = "FRED", from="2014-7-1", to="2015-7-1")

## Adjusting the fedFundsRate to daily return
fedFundsRate = (DFF[rownames(data.frame(SPY))]/365)/100

head(SPY)
head(fedFundsRate)
```

## Creating efficient frontier

```{r}
# Calculate mean and standard deviation for each stock
Mean.Sd.SP500.companies = cbind(
    sd = sapply(SP500.Industrials.Health.names, function(z) sd(diff(log(get(z)[,6])), na.rm=TRUE)),
    mean = sapply(SP500.Industrials.Health.names, function(z) mean(diff(log(get(z)[,6])), na.rm=TRUE))
)
# Calculate the mean and standard deviation for SPY (broad index)
Mean.Sd.SPY = c(sd(diff(log(SPY[,6])), na.rm=TRUE),mean(diff(log(SPY[,6])), na.rm=TRUE))
# Calculate the mean of Fed Funds
Mean.FedFunds = mean(fedFundsRate)
```

### Plotting the SPY companies on sd-mean diagram, with SPY and risk-free rate

```{r}
plot(Mean.Sd.SP500.companies, ylab = "Mean", xlab = "Standard Deviation (proxy for Risk)", main = "Efficient Frontier", pch = 19, xlim = c(0,.03))
points(Mean.Sd.SPY[1], Mean.Sd.SPY[2], col = "red", pch = 19)
points(0, Mean.FedFunds, col = "green", pch = 19)
lines(c(0,Mean.Sd.SPY[1]), c(mean(fedFundsRate), Mean.Sd.SPY[2]))
text(Mean.Sd.SP500.companies, labels = rownames(Mean.Sd.SP500.companies), cex = .5, pos = 2)
text(Mean.Sd.SPY[1], Mean.Sd.SPY[2], labels = "SPY", cex = .5, col = "red", pos = 2)
```

## Building CAPM model

```{r, warning = FALSE}
# Extracting the coefficients: Asset return - Risk free return = Beta(Market return - risk free return)
SP500.companies.betas = as.matrix(sapply(
        SP500.Industrials.Health.names, function(z) 
        lm(I(diff(log(get(z)[,6])) - fedFundsRate) ~ -1 + I(diff(log(SPY[,6])) - fedFundsRate))$coefficients
    )
)

rownames(SP500.companies.betas) = rownames(Mean.Sd.SP500.companies)
colnames(SP500.companies.betas) = "CAPM betas"

head(SP500.companies.betas)
```

### Plotting the Betas

```{r}
plot(
      SP500.companies.betas, Mean.Sd.SP500.companies[,2], pch = 19,
      main = "CAPM", xlim = c(0,2), ylab = "Expected Returns", xlab = "Beta", ylim = c(0, 0.001)
)
points(1, Mean.Sd.SPY[2], col = "red", pch = 19)
points(0, Mean.FedFunds, col = "green", pch = 19)
lines(c(0,1), c(Mean.FedFunds,Mean.Sd.SPY[2]))
text(SP500.companies.betas, Mean.Sd.SP500.companies[,2], labels = rownames(Mean.Sd.SP500.companies), cex = .5, pos = 2)
text(1, Mean.Sd.SPY[2], labels = "SPY", cex = .5, col = "red", pos = 2)
```

## Arbitrage Pricing Theory

```{r}

Stock.Portfolio.Returns = data.frame(matrix(NA, nrow = 251, ncol = length(SP500.Industrials.Health.names)))
colnames(Stock.Portfolio.Returns) = SP500.Industrials.Health.names
for (i in SP500.Industrials.Health.names){
  Stock.Portfolio.Returns[,i] = diff(log((get(i)[,6])))[-1,]
}

head(Stock.Portfolio.Returns)

#Returns for the indices
SPY.returns = as.matrix(diff(log(SPY[,6])))
XLV.returns = as.matrix(diff(log(XLV[,6])))
XLI.returns = as.matrix(diff(log(XLI[,6])))
```

### Step 1 Selection of factors

Running the PCA on portfolio

```{r}
Stock.Portfolio.Returns.PCA = princomp(Stock.Portfolio.Returns)
cumsum(Stock.Portfolio.Returns.PCA$sdev/sum(Stock.Portfolio.Returns.PCA$sdev))
```

From above table, we would need 13 components to describe the returns well (\>90%) but proceeding with 5 components since it explains more than 50% variance

```{r}
# Create 4 factors and 4 loadings
Stock.Portfolio.Returns.PCA.factors = as.matrix(Stock.Portfolio.Returns.PCA$scores[, 1:5])
Stock.Portfolio.Returns.PCA.loadings = Stock.Portfolio.Returns.PCA$loadings[, 1:5]
Stock.Portfolio.Returns.PCA.zero.loading = Stock.Portfolio.Returns.PCA$center
head(Stock.Portfolio.Returns.PCA.loadings)
# Plot the loadings
matplot(1:16, Stock.Portfolio.Returns.PCA.loadings, type="l", lty = 1)
abline(h = 0)
```

.

```{r}
plot(SPY.returns[-1], Stock.Portfolio.Returns.PCA.factors[,1], type = "p")
lm.fit.factor1 = lm(Stock.Portfolio.Returns.PCA.factors[,1] ~ SPY.returns[-1])
summary(lm.fit.factor1)
```

```{r}
plot((XLV.returns - XLI.returns)[-1], Stock.Portfolio.Returns.PCA.factors[,2], type = "p")
lm.fit.factor2 = lm(Stock.Portfolio.Returns.PCA.factors[,2] ~ I((XLV.returns-XLI.returns)[-1]))
summary(lm.fit.factor2)
```

The first model has a high R squared, while the second model does not.

```{r}
hist(lm.fit.factor1$residuals)
qqnorm(lm.fit.factor1$residuals)
qqline(lm.fit.factor1$residuals)

hist(lm.fit.factor2$residuals)
qqnorm(lm.fit.factor2$residuals)
qqline(lm.fit.factor2$residuals)
```

### Step 2 Estimating betas

```{r}
Stock.portfolio.betas = apply(
                             Stock.Portfolio.Returns, 2,
                             function(z) lm(z ~ Stock.Portfolio.Returns.PCA.factors[,1] + Stock.Portfolio.Returns.PCA.factors[,2])$coefficients
                        )
rownames(Stock.portfolio.betas) = c("alpha","factor1","factor2")
Stock.portfolio.betas = data.frame(t(Stock.portfolio.betas))
Stock.portfolio.betas
```

### Step 3. Estimating market price of risk

```{r}
Market.Prices.of.risk.fit = lm(I(alpha - Mean.FedFunds)~.-1, data = Stock.portfolio.betas)
summary(Market.Prices.of.risk.fit)

Market.Prices.of.risk = c(Mean.FedFunds, Market.Prices.of.risk.fit$coefficients)
Market.Prices.of.risk.fit$residuals
```

```{r}
plot(Market.Prices.of.risk.fit$residuals, type = "h", xaxt = "n", xlab = "Stock")
abline(h = 0)
axis(1, at = 1:16, labels = SP500.Industrials.Health.names)
```

When residual is positive it means that the mean return of the stock over the sample period E(R) is greater than predicted value. Looking at the residual plot, we see that CAT residuals are very negative, this suggests that the stock is overvalued but NOC and AET have high positive residuals, This suggests that These two stocks are undervalued.
