---
title: "Financial Analytics - Assignment Options"
author: "Jasmeet Singh Sandhu"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(RQuantLib)
library(NMOF)
```

```{r}
S = 810
sigma = 0.20
r = 0.05
d = 0.01
X = 800
t = 0.5

price = EuropeanOption(type = "call", underlying = S, strike = X,
                         dividendYield = d, riskFreeRate = r, maturity = t,
                         volatility = sigma)
price$value
```
```{r}
d1 = (log(S/X) + (r  + 0.5*(sigma^2))*t)/(sigma*sqrt(t))
d2 = d1 - (sigma*sqrt(t))

C = (S * pnorm(d1, 0, 1)) - ((X*exp(-r * t))*pnorm(d2,0,1))

cbind(calc_value = price$value, manual_value = C)
```
The values are not same

```{r}
div_yield_0 = EuropeanOption(type = "call", underlying = S, strike = X,
                         dividendYield = 0, riskFreeRate = r, maturity = t,
                         volatility = sigma)
cbind(calc_value = price$value, manual_value = C, calc_value_div_0 = div_yield_0$value)
```
Using div_yield = 0, we get same results = 61.25612


Higher dividend yields result in lower call prices. 
  - Account for dividend yield by using it to discount the Underlying price in the C equation.
  - C = S0 x (exp(-YT)) x N(d1) - (K x exp(-RT))N(d2) where Y is the dividend yield. 

Account for it in the equation for d1.
- d1 = (ln(S0/k) + (R - Y + 0.5*sigma^2) x T)/(sigma x sqrt(T))

However, the equation for d2 remains the same
- d2 = d1 - (sigma x sqrt(T))

```{r}
parity = price$value + (X*exp(-r*t)) - (S*exp(-d*t))
parity

put_price = EuropeanOption(type = "put", underlying = S, strike = X,
                         dividendYield = d, riskFreeRate = r, maturity = t,
                         volatility = sigma)
put_price$value
```
################################################ OPTIONS HEDGING ################################################

```{r}
datapath = 'C:\\Users\\jasme\\Downloads\\UChicago\\Quarter2\\Financial Analytics\\Week9'
df = read.csv(paste(datapath,'assignments_Fonancial Analytics (32001)_Lecture 7_OptionsPrices.CSV', sep='\\'),header = TRUE, row.names = 1, sep = ",")
head(df)
```

```{r}
calculateImpliedVolatility = function(dataRow){
  myExpiry = dataRow[1]/365
  myRate = dataRow[2]
  myUnderlying = dataRow[3]
  myPutStrikes = dataRow[4:8]
  myCallStrikes = dataRow[9:14]
  myPutPrices = dataRow[15:19]
  myCallPrices = dataRow[20:25]
  impVolPuts = apply(rbind(myPutPrices, myPutStrikes),2, 
                    function(z) EuropeanOptionImpliedVolatility("put", 
                                                                value=z[1],
                                                                underlying=myUnderlying, 
                                                                strike=z[2],
                                                                dividendYield=myRate,
                                                                riskFreeRate=myRate,
                                                                maturity=myExpiry,
                                                                volatility=.3))
  impVolCalls=apply(rbind(myCallPrices,myCallStrikes),2,
                     function(z) EuropeanOptionImpliedVolatility("call",
                                                                 value=z[1],
                                                                 underlying=myUnderlying,
                                                                 strike=z[2],
                                                                 dividendYield=myRate,
                                                                 riskFreeRate=myRate,
                                                                 maturity=myExpiry,
                                                                 volatility=.3))
  Result = cbind(Strikes=c(myPutStrikes,myCallStrikes),Volatilities=c(impVolPuts,impVolCalls))
  rownames(Result)=c(rep("P",5),rep("C",6))
  Result
}
```


```{r}
Row1Options = as.numeric(df[2,])
V.1 = calculateImpliedVolatility(as.numeric(Row1Options))
V.1
```
```{r}
put1425.impliedVolatilities = as.vector(apply(df, 1, 
                                               function (z) EuropeanOptionImpliedVolatility("put", value=z[26], 
                                                                                            underlying=z[3], strike=1425, 
                                                                                            dividendYield = z[2], 
                                                                                            riskFreeRate = z[2], 
                                                                                            maturity = z[1]/365, 
                                                                                            volatility = 0.3)))*100
currentPutValue = EuropeanOption(type = "put", underlying = df[1,3], strike = 1425, 
                                  maturity = df[1,1]/365, riskFreeRate = df[1,2], 
                                  volatility = put1425.impliedVolatilities[1]/100 , dividendYield = df[1,2])
currentPutValue
```
```{r}
putPNL.delta = as.numeric(currentPutValue)[2]*(df[2,3]-df[1,3])
putPNL.gamma= 0.5*as.numeric(currentPutValue)[3]*(df[2,3]-df[1,3])^2
putPNL.vega=as.numeric(currentPutValue)[4]/100*(put1425.impliedVolatilities[2]-put1425.impliedVolatilities[1])
(putPNL.estimated=c(delta=putPNL.delta,gamma=putPNL.gamma,vega=putPNL.vega))
```
```{r}
Row1Options = as.numeric(df[1,])
V.1 = calculateImpliedVolatility(as.numeric(Row1Options))
V.1
```

```{r}
plot(V.1, type="b")
```

```{r}
premia = rep(NA,11)
for (i in 1:11){
  premia[i] = unlist(EuropeanOption(type=ifelse(rownames(V.1)[i]=="P", "put", "call"),
                                     underlying = Row1Options[3],
                                     strike = V.1[i,1],
                                     dividendYield = Row1Options[2],
                                     riskFreeRate = Row1Options[2],
                                     maturity = Row1Options[1]/365,
                                     volatility = V.1[i,2]))[1]
}
checkPremia = cbind(strikes=Row1Options[4:14],
                     calculatedOptionPremia=unlist(premia),
                     originalOptionsPremia=Row1Options[15:25],
                     differences=unlist(premia)-Row1Options[15:25])
rownames(checkPremia) = colnames(df)[15:25]
checkPremia
```

```{r}
df[1:3,3]
```
```{r}
V.2 = calculateImpliedVolatility(as.numeric(df[2,]))
V.3 = calculateImpliedVolatility(as.numeric(df[3,]))
xBounds = c(min(V.1[,1],V.2[,1],V.3[,1]), max(V.1[,1],V.2[,1],V.3[,1]))
plotData.Strikes = seq(from=xBounds[1], to=xBounds[2], by =5)
plotData = rep(NA, 3*length(plotData.Strikes))
dim(plotData) = c(length(plotData.Strikes),3)
(plotData = cbind(plotData.Strikes,plotData)) 
```
```{r}
plotData[match(V.1[,1],plotData.Strikes),2] = V.1[,2]
plotData[match(V.2[,1],plotData.Strikes),3]=V.2[,2]
plotData[match(V.3[,1],plotData.Strikes),4]=V.3[,2]
plotData
```

```{r}
(volatilitiesATM = as.vector(apply((df), 1, function(z) calculateImpliedVolatility(z)[6,2]))*100)
```

```{r}
put1425.impliedVolatilities = as.vector(apply(df, 1, function (z) EuropeanOptionImpliedVolatility("put", 
                                                                                                  value=z[26], 
                                                                                                  underlying=z[3], 
                                                                                                  strike=1425, 
                                                                                                  dividendYield = z[2], 
                                                                                                  riskFreeRate = z[2], 
                                                                                                  maturity = z[1]/365, 
                                                                                                  volatility = 0.3)))*100
```

```{r}
df2 = cbind(df, put1425.impliedVolatilities)
head(df2)
put1425History = c()
put1425History = as.vector(apply(df2, 1, function(z) EuropeanOption(type="put",
                                     underlying = z[3],
                                     strike = 1425,
                                     dividendYield = z[2],
                                     riskFreeRate = z[2],
                                     maturity = z[1]/365,
                                     volatility = z[27]/100)$value))
```

```{r}
currentPutValue = EuropeanOption(type="put",underlying = df[1,3], strike = 1425, maturity = df[1,1]/365, riskFreeRate = df[1,2], volatility =put1425.impliedVolatilities[1]/100 , dividendYield = df[1,2])
currentPutValue
```
```{r}
VectorizedEIV  =  Vectorize(EuropeanOptionImpliedVolatility)

calculateImpliedVolatility  =  function(optionsData) {
  myExpiry  =  optionsData$Expiry / 365
  myRate  =  optionsData$OptRate
  myUnderlying  =  optionsData$UndPr
  myPutStrike  =  rep(1425, nrow(optionsData)) # set the strike to 1425
  myPutPrice  =  optionsData$P1425 # get the price from the 26th column
  impVolPut  =  VectorizedEIV("put",
                             value = myPutPrice,
                             underlying = myUnderlying,
                             strike = myPutStrike,
                             dividendYield = myRate,
                             riskFreeRate = myRate,
                             maturity = myExpiry,
                             volatility = 0.3) # set the initial volatility guess
  Result  =  data.frame(Volatility = as.numeric(impVolPut)) # create a dataframe with the results
  rownames(Result)  =  rownames(optionsData)
  Result
}


impliedVolPut1425  =  calculateImpliedVolatility(df)
head(impliedVolPut1425)
```

```{r}
VectorizedEO = Vectorize(EuropeanOption)
optionP1425 = t(VectorizedEO(type=rep("put", nrow(df)),
                      underlying=df$UndPr, 
                      strike=rep(1425,nrow(df)), 
                      dividendYield=df$OptRate,
                      riskFreeRate=df$OptRate, 
                      maturity=df$Expiry/365, 
                      volatility=impliedVolPut1425$Volatility))
optionDF = data.frame(value=as.numeric(optionP1425[,1]),
                       delta=as.numeric(optionP1425[,2]),
                       gamma=as.numeric(optionP1425[,3]),
                       vega=as.numeric(optionP1425[,4]),
                       theta=as.numeric(optionP1425[,5]),
                       rho=as.numeric(optionP1425[,6]),
                       divRho=as.numeric(optionP1425[,7]))
rownames(optionP1425) = rownames(df)
rownames(optionDF) = rownames(df)
head(optionDF)
```

```{r}
putPNL.delta = as.numeric(optionP1425[,2])*(diff(df$UndPr))
putPNL.gamma = 0.5*as.numeric(optionP1425[,3])*(diff(df$UndPr))^2
putPNL.vega = (as.numeric(optionP1425[,4])/100*(diff(impliedVolPut1425$Volatility)))*100
putPNL.estimatedTotal = putPNL.delta+putPNL.gamma+putPNL.vega
putPNL.actual = diff(df$P1425)
```
```{r}
myResults = data.frame(estimated=putPNL.estimatedTotal[-length(putPNL.estimatedTotal)],actual=putPNL.actual,delta=putPNL.delta[-length(putPNL.estimatedTotal)],gamma=putPNL.gamma[-length(putPNL.estimatedTotal)], vega=putPNL.vega[-length(putPNL.estimatedTotal)])
rownames(myResults) = rownames(df[-1,])
# myResults = myResults[1:5,]
myResults
```