---
title: "Financial Analytics Assignment 5"
author: "Jasmeet Singh Sandhu"
date: "`r Sys.Date()`"
output: html_document
---
# Analysis of Moody's Bond Yields

```{r, warning=FALSE, include=FALSE }
library(tseries)
library(forecast)
library(TSA)
library(urca)
# install.packages("fArma", repos="http://R-Forge.R-project.org")
library(fArma)
```

```{r}
datapath <- 'C:\\Users\\jasme\\Downloads\\UChicago\\Quarter2\\Financial Analytics\\Week5'
df = read.csv(file = paste(datapath, "assignments_Fonancial Analytics (32001)_Lecture 5_MYieldsData.csv", sep = '\\'))
aaa = df[,2]
baa = df[,3]
```

```{r}
# Plot the two time series
plot(aaa, col="red", type="l", main="AAA and BAA rates")
lines(baa, col="blue")
```

```{r}
plot(aaa, baa, col = 'blue', xlab="AAA Yields", ylab="BAA Yields", main="Scatter plot of AAA and BAA")
```
The series look co-related but when rates are low, there are deviations but when rates are high, the series are corelated.


```{r}
linear = lm(baa~aaa)
summary(linear)
```

```{r}
qqnorm(linear$residuals, col = 'blue')
acf(linear$residuals, col = 'blue',)
Box.test(linear$residuals, lag=12, type="Ljung")
```
The residuals are not normally distributed, hence not a good model

```{r}
diff.aaa = diff(aaa)
diff.baa = diff(baa)
plot(diff.aaa, diff.baa, col = 'blue', main="Scatterplot of yield differences")
```


```{r}
diff.linear = lm(diff.baa~diff.aaa-1)
summary(diff.linear)

qqnorm(diff.linear$residuals, col = 'blue')
acf(diff.linear$residuals, col = 'blue')
Box.test(diff.linear$residuals, lag=12, type="Ljung")
```

```{r}
dlr.ts = ts(diff.linear$residuals, frequency = 12)
auto.arima(dlr.ts, seasonal=T)
```
auto.arima suggesta ARMA(2,2) 

```{r}
arma22 = armaFit(~arma(2,2), data=dlr.ts)
summary(arma22, which="all")
```


```{r}
## Model = phi1*xt1 + phi2*xt2 + et + theta1*et1+theta2*et2
arma22.res = residuals(arma22) 
et2 = arma22.res
head(et2)

et1 = et2[-1]
head(et1)

et = et1[-1]
head(et)
```

```{r}
xt2 = dlr.ts[-length(dlr.ts)]
xt1 = dlr.ts[-1]
phi1 = arma22@fit$coef[1]
phi2 = arma22@fit$coef[2]
theta1 = arma22@fit$coef[3]
theta2 = arma22@fit$coef[4]

arma_model = phi1*xt1 + phi2*xt2 + et + theta1*et1+theta2*et2
head(arma_model)
```

forecast one step ahead for BAA

```{r}
x_BAA = baa[-length(aaa)]

forecast = x_BAA + diff.linear$coefficients * diff.aaa + arma_model

matplot(cbind(baa[-1],forecast),type = "l",col = c("red","blue"),main= "3 year rate and forecast", ylab="BAA yield and forecast")
legend("topright", c("BAA rates","forecasts"), lwd=2,col = c("red","blue"), bty="n")
```

```{r}
diff.aaa.2 = diff.aaa[-1]

diff.forecast = diff(forecast)

cr = cbind(diff.forecast, diff.aaa.2)

plot(
  cr[,1], cr[,2], col = "blue", 
  main = "Differences of forecasted BAA vs Differences of AAA Yields",
  xlab="Difference of forecasts of BAA Rate",ylab="Differences of AAA Rate"
)
```

## Cointegration

```{r}
data <- cbind(aaa, baa)

cajo <- ca.jo(data, ecdet = "none", type="eigen", K=2, spec ="longrun")
summary(cajo)
```

```{r}
plotres(cajo)
```

```{r}
cajo@teststat
```

```{r}
cajo@cval
```

```{r}
barplot(cajo@cval[1,],main = "Johansen test h<=1",col = "red")
abline(h=cajo@teststat[1], col="blue")
legend("topleft", c("critical values","test statistics"), lwd=2,col = c("red","blue"), bty="n")
```

```{r}
barplot(cajo@cval[2,],main = "Johansen test h=0",col = "red", ylim=range(0:25))
abline(h=cajo@teststat[2], col="blue")
legend("topleft", c("critical values","test statistics"), lwd=2,col = c("red","blue"), bty="n")
```
From this we can conclude that the series are cointegrated even at 1% significance level.

```{r}
(a_1 <- cajo@V[,1])
```

```{r}
z_t1= data %*% a_1
matplot(z_t1,type ="l", main = "z(1,t)=a1'x(t)", col = "blue")
```

```{r}
zar <-ar(z_t1,  aic = TRUE,method = "yule-walker")
zar$order
```

```{r}
armaRoots(zar$ar,lwd = 8, n.plot = 400, digits = 8)
```

```{r}
a_2<- cajo@V[,2]
z_t2= data %*% a_2
matplot(z_t2,type ="l", main = "z(2,t)=a2'x(t)", col = "blue")
```

### Predicting using Cointegration model

```{r}
(mu <- cajo@GAMMA[,1])
(PI <- cajo@PI)
(Gamma <- cajo@GAMMA[,2:3])
dX_1 <- cajo@Z0
head(dX_1)
X_2 <- cajo@ZK
head(X_2)
```

```{r}
deltaX_t_1 <- Gamma %*% t(dX_1) + PI %*%t(X_2) 
deltaX_t_1<-apply(deltaX_t_1,2,"+",mu)
nrowsdata <- dim(data)[1]
data_t_2 = data[3:nrowsdata,]
deltaX_t_1 <- t(deltaX_t_1)
forecastX <- data_t_2+deltaX_t_1
faaa = cbind(aaa[3:length(aaa)],forecastX[,1])
matplot(faaa,col =c("red","blue"),type="l",main = "AAA Yield and prediction")
legend("topright", c("AAA yield","prediction"), lwd=2,col = c("red","blue"), bty="n")
```

```{r}
fbaa = cbind(baa[3:length(baa)],forecastX[,2])
matplot(fbaa,col =c("red","blue"),type="l",main = "BAA yields and prediction")
legend("topright", c("BAA yield","prediction"), lwd=2,col = c("red", "blue"), bty="n")
```

```{r}
dfaaa = diff(faaa)
dfbaa = diff(fbaa)
plot(dfaaa,dfbaa,col ="blue", main = "Scatter plot for change of prediction for AAA and BAA yields",
     xlab="Differenced forecasts of AAA Yields", ylab="Differenced forecasts of BAA Yields")
```

```{r}
cerrorA = aaa[3:length(aaa)]-forecastX[,1]
matplot(cerrorA,col = 'blue', main = "Residual Prediction of AAA Yield", type = "l")
```

```{r}
cerrorB = baa[3:length(baa)]-forecastX[,2]
matplot(cerrorB,col = 'blue', main = "Residual Prediction of BAA Yield",type = "l")
```

```{r}
plot(cerrorA,cerrorB,col ="blue", main = "Scatter plot for residual prediction for AAA and BAA yields")
```

```{r}
cor(cbind(cerrorA, cerrorB))
```


Compare the errors of the regression model with the cointegration model

```{r}
linreg.errors = baa[-1] - forecast
errors = cbind(linreg.errors[-1], cerrorB)
matplot(errors,type ="l",col = c("orange","blue"),main = "BAA Yield Residual Regression and Cointegration Model")
legend("topright", c("regression errors","cointegration errors"), lwd=2,col = c("orange","blue"), bty="n")
```

The variance level of cointegration errors is lower than for errors of the regression model.

Check how errors of the two models are related to each other

```{r}
plot(errors[,1],errors[,2],col = "blue", 
     main = "Scatter Plot of Regression model Errors vs Cointegration errors",
     xlab="Regression Model Errors", ylab="Cointegration Model Errors")
```
The model errors look uncorrelated.
The Co integration performs good job of forecasting when comparing the error variance.
